GPU-friendly Stroke Expansion

### 2024-06-14 paper.tex

Correct stroke rendering also relies on accurate approximation of the parallel curves of the input paths. Determining the parallel curve of a cubic Bézier is not analytically tractable, as it is a tenth order algebraic formula[\citenum{Farouki1990}]. Thus, all practical stroke rendering techniques employ an approximation, with error tolerance as a tunable parameter, typically representing the maximum allowable Fréchet distance between the true curve and its approximation. A typical value is 0.25 device pixel, as that is close to the threshold of visibility. A correct algorithm will produce an approximate curve within the error tolerance, but an efficient algorithm will not subdivide significantly more than is necessary to meet the error bound. Some specifications for vector graphics, including PDF[\citenum{PDF2008}], specify the error tolerance explicitly.

The subproblem of approximating parallel curves has spawned extensive literature, none of which is entirely satisfying, especially when it comes to algorithms that can be efficiently evaluated on GPU. An example of a reasonably good algorithm that computes flattened parallel curves is \citet{Yzerman2020}, as used in the Blend2D rendering library. For producing curved outlines, the Tiller and Hanson[\citenum{Tiller1984}] algorithm is commonly implemented and cited, but its performance is quite poor when applied to cubic Béziers. The quadratic Bézier version is adequate, and forms the basis of the algorithm in \citet{Nehab2020}. A variant is also used in Skia[\citenum{Skia}]; instead of lowering the cubic Bézier to a quadratic approximation and then computing the parallel curve, Skia approximates the parallel curve directly with a quadratic Bézier, then measures the error to determine whether further subdivision is necessary.

@@

\begin{algorithm}
\SetKwProg{Fn}{fn}{}{end}
\SetKwBlock{Loop}{loop}{end}
\Fn{flatten\_cubic(cubic: Cubic, offset: f32)}{
    TOL $\gets$ Euler spiral fit tolerance\;
    \tcp{'t0\_u' and 'dt' track the recursion stack}
    t0\_u $\gets$ 0\;
    dt $\gets$ 1.0\;
    last\_p $\gets$ cubic.p0\;
    last\_q $\gets$ cubic.p1 - cubic.p0\;
    \Loop{
        t0 $\gets$ f32(t0\_u) * dt\;
        \If{t0 == 1.0}{break\;}
        t1 $\gets$ t0 + dt\;
        (next\_p, next\_q) $\gets$ eval\_cubic\_and\_deriv(cubic, t1)\;
        \tcp{Estimate error of geometric}
        \tcp{Hermite interpolation to Euler spiral}
        \nosemic(th0, th1, chord\_len, err) $\gets$ fit\_cubic\_to\_es(\;
        \pushline\dosemic last\_p, next\_p, last\_q, next\_q, dt)\;
        \popline \eIf{err <= TOL}{
            es $\gets$ euler\_segment(th0, th1, last\_p, next\_p)\;
            \tcp{Find location of any cusp in outline.}
            \tcp{If $0 \leq t < 1$, draw evolute at 't'.}
            cusp0 $\gets$ es\_curvature(es, 0) * offset + chord\_len\;
            cusp1 $\gets$ es\_curvature(es, 1) * offset + chord\_len\;
            t $\gets$ cusp0 * cusp1 >= 0 ? 1 : cusp0 / (cusp0 - cusp1)\;
            \If{cusp0 >= 0}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(0, t))\;
            }
            \If{cusp0 < 0 OR t < 1}{
                evolute\_start(es, offset)\;
                range $\gets$ cusp0 >= 0 ? vec2(t, 1) : vec2(0, t)\;
                es\_seg\_flatten\_evolute(es, range)\;
                es\_seg\_flatten\_offset\_reverse(es, offset, range)\;
            }
            \If{cusp0 < 0 AND t < 1}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(t, 1))\;
            }
            \tcp{Pop the stack}
            t0\_u $\gets$ t0\_u + 1\;
            shift $\gets$ countTrailingZeros(t0\_u)\;
            t0\_u $\gets$ t0\_u >> shift;
            dt $\gets$ dt * f32(1 << shift)\;
            last\_p $\gets$ next\_p\;
            last\_q $\gets$ next\_q\;
        }{
            \tcp{Error is above the tolerance.}
            \tcp{Push the stack; Continue subdivision}
            t0\_u $\gets$ t0\_u * 2\;
            dt $\gets$ dt * 0.5\;
        }
    }
}
\caption{Outline of the iterative cubic Bézier flattening routine invoked in our compute shader. A working WGSL implementation is provided in the supplemental materials.}
\label{alg:flatten-cubic}
\end{algorithm}

All input segments first get converted to a cubic Bézier by degree elevation. Our flattening routine (see Algorithm~\ref{alg:flatten-cubic}) starts by computing the error estimate for geometric Hermite interpolation from the cubic Bézier to an Euler spiral segment. If the error exceeds the desired tolerance threshold, the cubic segment is iteratively subdivided in half and a new error estimate is computed. Once the error satisfies the tolerance, the Euler spiral segment is directly flattened to a polyline (or, alternately, converted to a sequence of arc segments) by invoking a subroutine that is parameterized by the stroke halfwidth (see \texttt{es\_seg\_flatten\_offset} in Algoritm~\ref{alg:flatten-cubic}). For a regular fill, this parameter is set to $0$. For strokes, the sign of the parameter determines the position of the parallel curve relative to the source segment as well as the winding direction of the output lines.

This procedure is sufficient to generate weakly correct stroke outlines. If the outline contains a cusp -- which can be determined by evaluating the curvature variation along the Euler spiral segment -- we output an evolute patch (as shown in Figure~\ref{fig:evolutes}) in order to achieve strong correctness. A selection of renderings produced by our algorithm that demonstrate both strong and weak correctness are shown in Figure~\ref{fig:stroke_stills}.

The output of the compute shader is a collection of line segments that gets stored in a GPU storage buffer. Every element in the output contains the viewport coordinates of the two endpoints of a single line segment. A property of the data-parallel structure is that the line segments are unordered. Because the segments are processed in parallel, ordering cannot be guaranteed across path segments. We refer to this output data structure as \emph{line soup}.

An application may want to sort the line soup, depending on the requirements of the chosen rendering algorithm. For instance, our own renderer employs tile-based scanline rasterization in a compute shader, with a pipeline architecture that resembles \emph{cudaraster}[\citenum{Laine2011}] and \emph{MPVG}[\citenum{Ganacim2014}]. In this design, the line soup elements are spatially sorted by subsequent compute stages into a grid of 16x16 pixel tiles. The final compute stage of our renderer (called \emph{fine rasterizer}) operates at tile granularity and computes pixel coverage for filled polygons outlined by the line soup (Figure~\ref{fig:pipeline} illustrates the overall structure of our renderer, the details of which are beyond the scope of this paper).

\begin{figure*}
    \includegraphics[scale=0.275]{pipeline}
    \caption{We integrated our curve flattening shader into our own renderer implemented entirely in compute shaders. The flattening stage (described in Section~\ref{section:gpu-impl}) converts stroked paths to filled outlines and flattens them into an unordered soup of line segments. Our renderer spatially sorts the line soup into tiles before computing pixel coverage using a highly parallel scanline rasterizer.}
    \Description{Our GPU pipeline architecture in broad strokes}
    \label{fig:pipeline}
\end{figure*}

We also implemented a variant that outputs circular arc segments instead of lines, in which the output entries contain the arc endpoints in addition to its curvature. For both primitive types, our renderer also outputs a 32-bit \emph{path ID} used by downstream pipeline stages to identify the path that a segment belongs to, in order to retrieve the correct paint (solid or gradient color) for rendering.

@@

We authored two additional test scenes in order to stress the GPU with a higher workload. The first is a very large stroked path with ~500,000 individually styled dashed segments which we adapted from Skia's open-source test corpus. We used two variants of this scene with \emph{butt} and \emph{round} cap styles. For the second scene we adapted the \emph{Canvas Paths} test case from MotionMark[\citenum{MotionMark}], which renders a large number of randomly generated stroked line, quadratic, and cubic Bézier paths. We made two versions with different sizes: mmark-70k and mmark-120k with 70,000 and 120,000 path segments, respectively. Table~\ref{table:bump-counts} shows the payload sizes of our largest test scenes.

\begin{table}
    \caption{Comparison of input and output segment counts in the more intensive GPU test cases. \emph{Segments} are input path segments. \emph{Lines} and \emph{Arcs} are the output primitives.}
    \begin{tabular}{|l|c|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Input Segments} & \textbf{Lines} & \textbf{Arcs} \\
    \hline
        waves.svg & 13,308 & 406,059 & 111,965 \\
        mmark-70k & 70,000 & 1,039,899 & 500,351 \\
        mmark-120k & 120,000 & 1,784,896 & 858,162 \\
        long dash (butt) & 503,304 & 1,060,476 & 1,060,476 \\
        long dash (round) & 503,304 & 2,013,216 & 1,060,476 \\
    \hline
    \end{tabular}
    \label{table:bump-counts}
\end{table}

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \citet{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in about ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \citet{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

@@

Figures~\ref{fig:nehab-gpu-timings} and \ref{fig:test-scenes-gpu-timings} show the execution times for the two data sets. The entire \citet{Nehab2020} timings set adds up to less than 1 ms on all GPUs except the mobile unit. Our algorithm is ~14x slower on the Mali-G78 compared to the M1 Max but it is still capable of processing \emph{waves} in 3 ms. We observed that the performance of the kernel scales well with increasing GPU ability even at very large workloads, as demonstrated by the order-of-magnitude difference in run time between the mobile, laptop, and high-end desktop form factors we tested. We also confirmed that outputting arcs instead of lines can lead to a 2x decrease in execution time (particularly in scenes with high curve content, like \emph{waves} and \emph{mmark}) as predicted by the lower overall primitive count.

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel. The CPU encoding times (prior to the GPU prefix sums) for some of the large scenes is shown in Table~\ref{table:bump-estimate-cost}. Our implementation applies dashing on the CPU. It is conceptually fairly straightforwad to implement on GPU; the core algorithm is prefix sum applied to segment length, and the Euler spiral representation is particularly well suited. However, there are complex implementation tradeoffs, so the dash style applied to the \emph{long dash} scene had to be processed on the CPU and dashes were sequentially encoded as individual strokes. Including CPU-side dashing, the time to encode \emph{long dash} on the Apple M1 Max was approximately 25 ms on average, while the average time to encode the path after dashing was 4.14 ms.

@@

We augmented our encoding logic to compute a conservative estimate of the required output buffer size based on an upper bound on the number of line primitives per input path segment. We found that Wang's formula[\citenum{Goldman2003}] works well in practice, as it is relatively inexpensive to evaluate, though it results in a higher memory estimate than required (see Table~\ref{table:bump-estimate-mem}). For parallel curves, Wang's formula does not guarantee an accurate upper bound on the required subdivisions, which we worked around by inflating the estimate for the source segment by a scale factor derived from the linewidth. We observed that enabling estimation increases our CPU pre-processing time by 2--3x on the largest scenes but the overall impact is negligible when the scene size is modest.

@@

\begin{table}
    \caption{The estimated and actual \emph{line soup} buffer sizes for scenes rendered at 2088x1600 resolution. Our estimation uses Wang's formula combined with a scaling heuristic based on the linewidth. In the worst case, our buffer size estimator overshoots by a factor of 4.}
    \begin{tabular}{|l|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & Actual Size & Estimated Size \\
    \hline
        roads.svg  & 958.78 KB & 2.45 MB    \\
        waves.svg  & 9.37 MB   & 10.44 MB   \\
        mmark-70k  & 23.81 MB  & 100.79 MB  \\
        mmark-120k & 40.86 MB  & 173.56 MB  \\
    \hline
    \end{tabular}
    \label{table:bump-estimate-mem}
\end{table}

\begin{figure*}
    \centering
    \includegraphics[scale=0.2]{test_renderings}
    \caption{Renderings of some of our test scenes created using our algorithm and associated frame rates. All images were rendered on an Apple M1 Max at 2088x1600 resolution using our GPU pipeline. \textbf{(a)} \emph{waves.svg} from the Nehab2020 timings data set (340fps); \textbf{(b)} The \emph{mmark-120k} scene with 120,000 input segments (72fps); \textbf{(c)} The \emph{long path} scene with ~500k input segments and round caps (40fps with CPU-side dashing - the GPU time to render the scene is 9ms); \textbf{(d)}; A simple test scene showcasing various cap and join styles (950fps); \textbf{(e)} The famous \emph{Ghostscript Tiger} containing both stroked and filled paths (810fps).}
    \Description{Renderings of the test scenes}
    \label{fig:renderings}
\end{figure*}

>>>>

Correct stroke rendering also relies on accurate approximation of the parallel curves of the input paths. Determining the parallel curve of a cubic Bézier is not analytically tractable, as it is a tenth order algebraic formula~[\citenum{Farouki1990}]. Thus, all practical stroke rendering techniques employ an approximation, with error tolerance as a tunable parameter, typically representing the maximum allowable Fréchet distance between the true curve and its approximation. A typical value is 0.25 device pixel, as that is close to the threshold of visibility. A correct algorithm will produce an approximate curve within the error tolerance, but an efficient algorithm will not subdivide significantly more than is necessary to meet the error bound. Some specifications for vector graphics, including PDF~[\citenum{PDF2008}], specify the error tolerance explicitly.

The subproblem of approximating parallel curves has spawned extensive literature, none of which is entirely satisfying, especially when it comes to algorithms that can be efficiently evaluated on GPU. An example of a reasonably good algorithm that computes flattened parallel curves is \citet{Yzerman2020}, as used in the Blend2D rendering library. For producing curved outlines, the Tiller and Hanson~[\citenum{Tiller1984}] algorithm is commonly implemented and cited, but its performance is quite poor when applied to cubic Béziers. The quadratic Bézier version is adequate, and forms the basis of the algorithm in \citet{Nehab2020}. A variant is also used in Skia~[\citenum{Skia}]; instead of lowering the cubic Bézier to a quadratic approximation and then computing the parallel curve, Skia approximates the parallel curve directly with a quadratic Bézier, then measures the error to determine whether further subdivision is necessary.

@@

\begin{algorithm}
\SetKwProg{Fn}{fn}{}{end}
\SetKwBlock{Loop}{loop}{end}
\Fn{flatten\_cubic(cubic: Cubic, offset: f32)}{
    TOL $\gets$ Euler spiral fit tolerance\;
    \tcp{`t0\_u' and `dt' track the recursion stack}
    t0\_u $\gets$ 0\;
    dt $\gets$ 1.0\;
    last\_p $\gets$ cubic.p0\;
    last\_q $\gets$ cubic.p1 - cubic.p0\;
    \Loop{
        t0 $\gets$ f32(t0\_u) * dt\;
        \If{t0 == 1.0}{break\;}
        t1 $\gets$ t0 + dt\;
        (next\_p, next\_q) $\gets$ eval\_cubic\_and\_deriv(cubic, t1)\;
        \tcp{Estimate error of geometric}
        \tcp{Hermite interpolation to Euler spiral}
        \nosemic(th0, th1, chord\_len, err) $\gets$ fit\_cubic\_to\_es(\;
        \pushline\dosemic last\_p, next\_p, last\_q, next\_q, dt)\;
        \popline \eIf{err <= TOL}{
            es $\gets$ euler\_segment(th0, th1, last\_p, next\_p)\;
            \tcp{Find location of any cusp in outline.}
            \tcp{If $0 \leq t < 1$, draw evolute at `t'.}
            cusp0 $\gets$ es\_curvature(es, 0) * offset + chord\_len\;
            cusp1 $\gets$ es\_curvature(es, 1) * offset + chord\_len\;
            t $\gets$ cusp0 * cusp1 >= 0 ? 1 : cusp0 / (cusp0 - cusp1)\;
            \If{cusp0 >= 0}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(0, t))\;
            }
            \If{cusp0 < 0 OR t < 1}{
                evolute\_start(es, offset)\;
                range $\gets$ cusp0 >= 0 ? vec2(t, 1) : vec2(0, t)\;
                es\_seg\_flatten\_evolute(es, range)\;
                es\_seg\_flatten\_offset\_reverse(es, offset, range)\;
            }
            \If{cusp0 < 0 AND t < 1}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(t, 1))\;
            }
            \tcp{Pop the stack}
            t0\_u $\gets$ t0\_u + 1\;
            shift $\gets$ countTrailingZeros(t0\_u)\;
            t0\_u $\gets$ t0\_u $\gg$ shift\;
            dt $\gets$ dt * f32(1 $\ll$ shift)\;
            last\_p $\gets$ next\_p\;
            last\_q $\gets$ next\_q\;
        }{
            \tcp{Error is above the tolerance.}
            \tcp{Push the stack; Continue subdivision}
            t0\_u $\gets$ t0\_u * 2\;
            dt $\gets$ dt * 0.5\;
        }
    }
}
\caption{Outline of the iterative cubic Bézier flattening routine invoked in our compute shader. A working WGSL implementation is provided in the supplemental materials.}
\label{alg:flatten-cubic}
\end{algorithm}

All input segments first get converted to a cubic Bézier by degree elevation. Our flattening routine (see Algorithm~\ref{alg:flatten-cubic}) starts by computing the error estimate for geometric Hermite interpolation from the cubic Bézier to an Euler spiral segment. If the error exceeds the desired tolerance threshold, the cubic segment is iteratively subdivided in half and a new error estimate is computed. Once the error satisfies the tolerance, the Euler spiral segment is directly flattened to a polyline by invoking a subroutine that is parameterized by the stroke half-width (see \texttt{es\_seg\_flatten\_offset} in Algoritm~\ref{alg:flatten-cubic}). For a regular fill, this parameter is set to $0$. For strokes, the sign of the parameter determines the position of the parallel curve relative to the source segment as well as the winding direction of the output lines.

This procedure is sufficient to generate weakly correct stroke outlines. If the outline contains a cusp -- which can be determined by evaluating the curvature variation along the Euler spiral segment -- we output an evolute patch (as shown in Figure~\ref{fig:evolutes}) in order to achieve strong correctness. A selection of renderings produced by our algorithm that demonstrate both strong and weak correctness is shown in Figure~\ref{fig:stroke_stills}.

The output of the compute shader is a collection of line segments that gets stored in a GPU storage buffer. Every element in the output contains the viewport coordinates of the two endpoints of a single line segment. Threads reserve storage buffer regions for their output by incrementing an atomic index stored in GPU device memory. A property of the data-parallel structure is that the line segments are unordered. Because the segments are processed in parallel, ordering cannot be guaranteed across path segments. We refer to this output data structure as \emph{line soup}.

An application may want to sort the line soup, depending on the requirements of the chosen rendering algorithm. For instance, our own renderer employs tile-based scanline rasterization in a compute shader, with a pipeline architecture that resembles \emph{cudaraster}~[\citenum{Laine2011}] and \emph{MPVG}~[\citenum{Ganacim2014}]. In this design, the line soup elements are spatially sorted by subsequent compute stages into a grid of 16x16 pixel tiles. The final compute stage of our renderer (called \emph{fine rasterizer}) operates at tile granularity and computes pixel coverage for filled polygons outlined by the line soup (Figure~\ref{fig:pipeline} illustrates the overall structure of our renderer, the details of which are beyond the scope of this paper).

\begin{figure*}
    \includegraphics[scale=0.275]{pipeline}
    \caption{We integrated our curve flattening shader into our own renderer implemented entirely in compute shaders. The flattening stage (described in Section~\ref{section:gpu-impl}) converts stroked paths to filled outlines and flattens them to an unordered soup of line segments. Our renderer spatially sorts the line soup into tiles before computing pixel coverage using a highly parallel scanline rasterizer.}
    \Description{Our GPU pipeline architecture in broad strokes}
    \label{fig:pipeline}
\end{figure*}

We also implemented a variant that outputs circular arc segments instead of lines, in which the output entries store curvature in addition to the endpoints. For both primitive types, our renderer also outputs a 32-bit \emph{path ID} used by downstream pipeline stages to identify the path that a segment belongs to, in order to retrieve the correct paint (solid or gradient color) for rendering.

@@

We authored two additional test scenes in order stress the GPU with a higher workload. The first is a very large stroked path with \textasciitilde500,000 individually styled dashed segments which we adapted from Skia's open-source test corpus. We used two variants of this scene with \emph{butt} and \emph{round} cap styles. For the second scene we adapted the \emph{Canvas Paths} test case from MotionMark~[\citenum{MotionMark}], which renders a large number of randomly generated stroked line, quadratic, and cubic Bézier paths. We made two versions with different sizes: mmark-70k and mmark-120k with 70,000 and 120,000 path segments, respectively. Table~\ref{table:bump-counts} shows the payload sizes of our largest test scenes.

\begin{table}
    \caption{Comparison of input and output segment counts in the more intensive GPU test cases. \emph{Segments} are input path segments. \emph{Lines} and \emph{Arcs} are the output primitives.}
    \begin{tabular}{|l|c|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & \textbf{Input Segments} & \textbf{Lines} & \textbf{Arcs} \\
    \hline
        waves.svg & 13,308 & 475,855 & 181,229 \\
        mmark-70k & 70,000 & 1,577,705 & 1,162,117 \\
        mmark-120k & 120,000 & 2,709,013 & 1,994,946 \\
        long dash (butt) & 503,304 & 1,672,561 & 1,672,561 \\
        long dash (round) & 503,304 & 2,625,300 & 1,672,561 \\
    \hline
    \end{tabular}
    \label{table:bump-counts}
\end{table}

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \citet{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in \textasciitilde3 ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \citet{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

@@

Figures~\ref{fig:nehab-gpu-timings} and \ref{fig:test-scenes-gpu-timings} show the execution times for the two data sets. The entire \citet{Nehab2020} timings set adds up to less than 1 ms on all GPUs except the mobile unit. Our algorithm is \textasciitilde14$\times$ slower on the Mali-G78 compared to the M1 Max but it is still capable of processing \emph{waves} in 3~ms. We observed that the performance of the kernel scales well with increasing GPU ability even at very large workloads, as demonstrated by the order-of-magnitude difference in run time between the mobile, laptop, and high-end desktop form factors we tested. We also confirmed that outputting arcs instead of lines can lead to a 2$\times$ decrease in execution time (particularly in scenes with high curve content, like \emph{waves} and \emph{mmark}) as predicted by the lower overall primitive count.

@@

Figure~\ref{fig:tag-monoid-gpu-timings} shows the run time of the tag monoid parallel prefix sums with increasing input size. The prefix sums scale better at large input sizes compared to our stroke expansion kernel. This can be attributed to a lack of expensive computations and high control-flow uniformity in the tag monoid kernel. The CPU encoding times (prior to the GPU prefix sums) for some of the large scenes is shown in Table~\ref{table:bump-estimate-cost}. Our implementation applies dashing on the CPU. It is conceptually fairly straightforward to implement on GPU; the core algorithm is prefix sum applied to segment length, and the Euler spiral representation is particularly well suited. However, there are complex implementation tradeoffs, so the dash style applied to the \emph{long dash} scene had to be processed on the CPU and dashes were sequentially encoded as individual strokes. Including CPU-side dashing, the time to encode \emph{long dash} on the Apple M1 Max was approximately 25~ms on average, while the average time to encode the path after dashing was 4.14~ms.

@@

We augmented our encoding logic to compute a conservative estimate of the required output buffer size based on an upper bound on the number of line primitives per input path segment. We found that Wang's formula~[\citenum{Goldman2003}] works well in practice, as it is relatively inexpensive to evaluate, though it results in a higher memory estimate than required (see Table~\ref{table:bump-estimate-mem}). For parallel curves, Wang's formula does not guarantee an accurate upper bound on the required subdivisions, which we worked around by inflating the estimate for the source segment by a scale factor derived from the linewidth. We observed that enabling estimation increases our CPU pre-processing time by 2--3$\times$ on the largest scenes but the impact is overall negligible when the scene size is modest.

@@

\begin{table}
    \caption{The estimated and actual \emph{line soup} buffer sizes for scenes rendered at 2088x1600 resolution. Our estimation uses Wang's formula combined with a scaling heuristic based on the linewidth. The estimate is conservative and always overshoots the actual memory requirement.}
    \begin{tabular}{|l|c|c|}
    \hline
        \multicolumn{1}{|c|}{\textbf{Test}} & Actual Size & Estimated Size \\
    \hline
        roads.svg  & 2.76 MB   & 8.35 MB    \\
        waves.svg  & 10.89 MB  & 80.69 MB   \\
        mmark-70k  & 36.10 MB  & 103.06 MB  \\
        mmark-120k & 60.01 MB  & 177.33 MB  \\
    \hline
    \end{tabular}
    \label{table:bump-estimate-mem}
\end{table}

\begin{figure*}
    \centering
    \includegraphics[scale=0.2]{test_renderings}
    \caption{Renderings of some of our test scenes created using our algorithm. All images were rendered on an Apple M1 Max at 2088x1600 resolution using our GPU pipeline. \textbf{(a)} \emph{waves.svg} from the Nehab2020 timings data set (3.09~ms); \textbf{(b)} The \emph{mmark-120k} scene with 120,000 input segments (19.86~ms); \textbf{(c)} The \emph{long path} scene with \textasciitilde500k input segments and round caps (25~ms including CPU-side dashing -- the GPU time to render the scene is 9~ms); \textbf{(d)}; A simple test scene showcasing various cap and join styles (1.56~ms); \textbf{(e)} The famous \emph{Ghostscript Tiger} containing both stroked and filled paths (1.69~ms).}
    \Description{Renderings of the test scenes}
    \label{fig:renderings}
\end{figure*}
